The Calibration Reference Data System

### 2016-03-14 docs/paper/response.txt

1) box text file figures
2) improve screen shots

===================================================

Reviewer 2

General comment regarding databases and comparisions to other observatories.

A section at the end has been added that briefly describes what Spitzer, 
Chandra and Gemini use and notes some of the restrictions entailed in their
approaches (Chandra's is most similar in approach but simpler).

As far as the implications regarding databases go, we have not expanded on that.
This is mainly because some issues are arguably self evident (distributing
software that requires databases certainly does impose a significantly larger 
burden on those installing and maintaining such a system; we doubt that is 
controversial).

We also think it is nearly self evident that SQL does not always represent 
rules in a clear or easy to understand way (and many astronomers, the ones 
that ultimately must provide those rules, are not experts in SQL, particularly
complex SQL). We aren't saying it is impossible to use only SQL, just not 
ideal. Indeed, Gemini's experience found it to be too constricting (as we did).
Chandra doesn't use it at all, and it appears Spitzer's needs are not as complex.
This is probably aided by the fact that they generally process the data in 
large runs, i.e,. their updates are relatively infrequent as we understand 
it.  Besides, they sometimes have special processing after the query (e.g., 
interpolation between two reference files), so it isn't a pure database
solution either. Accomodating new needs requires new software changes.
To be fair, so does ours if a new kind of rule is required, but that
is much less likely, and much more likely to be in adding a new module
rather than changing existing code.

As to the focus on STScI needs. It would be nice if we had the resources 
to develop a system with a very broad reach and goals initially (well, perhaps,
many very ambitious projects die for being too ambitious!). But with the
reality of funding oversight, it's very difficult to do that. NASA gets 
quite nervous when they see work being done that isn't specifically done
for their mission. We do what we can to keep the project more broadly 
useful within these constraints. That's even true when writing papers
(i.e., can we justify spending much time doing a thorough survey of all
systems out there?).

@@

Section 2
We have reworded to be more explicit that all possible combinations of parameters
must be represented as distinct rows.

Section 3 point 2: "Could an 'is valid' boolean attribute in the database
help out?"
Not really. The file itself may be perfectly good, it is just associated with
the wrong value of parameters. Possbily if coupled with a resubmission with the
right values, but note that may mean many entries have to be so marked, and 
replicated. But as implemented, it didn't have that option. If it had been 
designed from the beginning to handle that, yes, it could have dealt with bad
submissions that way.

### 2016-03-14 docs/paper/crds.tex
updated all text except for pseudo-mode clarification
git-svn-id: https://aeon.stsci.edu/ssb/svn/crds_server/trunk@2248 37088bc5-bd3f-4c5b-b9de-119bb1e8f7e3

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

\usepackage{mdframed}
%% The amssymb package provides various useful mathematical symbols
% \usepackage{amssymb}
\usepackage{gensymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

We describe a software architecture and implementation for using rules to
determine which calibration files are appropriate for calibrating a
given observation. This new system, the Calibration Reference Data System
(CRDS), replaces what had been previously used for
the Hubble Space Telescope (HST) calibration pipelines, 
the Calibration Database System (CDBS). 
CRDS will be used for the James Webb Space
Telescope (JWST) calibration pipelines, and is currently
being used for HST calibration pipelines. CRDS can be
easily generalized for use in similar applications that need a rules-based
system for selecting the appropriate item for a given dataset; we give some
examples of such generalizations that will likely be used for JWST. The core
functionality of the Calibration Reference Data System is available under an
Open Source license.
\end{abstract}

@@

\item \textbf{Difficulty in undoing mistakes.}
A delivery of a file with mistaken
parameters could be difficult to undo. No reference file could be removed from
the database for various reasons. In a number of cases, particularly if the
date that it was targeted for was incorrect, it could seriously corrupt the
selection logic. To take the most common error case: Files often had a
``useafter'' date associated with them that indicated that it should be used for
a date after the specified one \textit{if no other file had a later date that was
still before the date of the observation}. If one submitted a file intended to
be used for data after 2000-01-01 but mistakenly provided
1999-01-01 then two new file submissions are required to correct the problem.
One to resubmit the new file with the correct date; the second to resubmit the file
intended to be used after 1999-01-01 so that its useafter date is one second after
the mistaken submission.

@@

\begin{figure*}
\begin{mdframed}
\begin{verbatim}
header = {
    'derived_from' : 'jwst_miri_dark_0007.rmap',
    'filekind' : 'DARK',
    'instrument' : 'MIRI',
    'mapping' : 'REFERENCE',
    'name' : 'jwst_miri_dark_0009.rmap',
    'observatory' : 'JWST',
    'parkey' : (('META.INSTRUMENT.DETECTOR', 'META.EXPOSURE.READPATT', 'META.SUBARRAY.NAME'),),
    'sha1sum' : 'bd99f4960096de9ad39a9ce506d5ff985847c4a3',
}

selector = Match({
    ('MIRIFULONG', 'FAST', 'FULL') : UseAfter({
        '2015-01-27 12:05:34' 'jwst_miri_dark_0029.fits',
        '2015-03-22 00:29:01' 'jwst_miri_dark_0032.fits',
        '2015-04-07 09:14:00' 'jwst_miri_dark_0047.fits',
        ...
        }),
    ('MIRIFUSHORT', 'SLOW', 'FULL') : UseAfter({
        '2015-01-25 12:06:34' 'jwst_miri_dark_0099.fits',
        '2015-03-21 00:28:01' 'jwst_miri_dark_0101.fits',
        '2015-04-08 09:13:00' 'jwst_miri_dark_0127.fits',
        ...
        }),
        ...
})
\end{verbatim}
\end{mdframed}
\caption{Example showing basic parameter matching rules coupled with date dependency.}
\label{fig:example1}
\end{figure*}

@@

\begin{figure*}
\begin{mdframed}
\begin{verbatim}
selector = SelectVersion({
   '<3.1':    'jwst_miri_flat_0065.fits',
   '<5':      'jwst_miri_flat_0073.fits',
   'default': 'jwst_miri_flat_123.fits',
})
\end{verbatim}
\end{mdframed}
\caption{Example showing use of the software version selection mechanism.}
\label{fig:example2}
\end{figure*}

@@

\begin{figure*}
\begin{mdframed}
\begin{verbatim}
selector = Bracket({
    1.2: "cref_flatfield_120.fits",
    1.5: "cref_flatfield_124.fits",
    5.0: "cref_flatfield_137.fits",
})
\end{verbatim}
\end{mdframed}
\caption{Example showing bracketing selection.}
\label{fig:example3}
\end{figure*}

@@

\begin{figure*}
\begin{mdframed}
\begin{verbatim}
selector = Match({
    ('WFC', 'N/A', 'BETWEEN 1220 1530', 'FR782N') :  "hst_wfc3_flatfile_0027.fits",
    ('WFC', 'N/A', 'BETWEEN 1530 1690', 'FR782N') :  "hst_wfc3_flatfile_0028.fits",
    ...
})
\end{verbatim}
\end{mdframed}
\caption{An example showing matching specified ranges.}
\label{fig:example4}
\end{figure*}

@@

\begin{figure*}
\begin{mdframed}
\begin{verbatim}
header = {
    'derived_from' : 'jwst_nircam_0015.imap',
    'instrument' : 'NIRCAM',
    'mapping' : 'INSTRUMENT',
    'name' : 'jwst_nircam_0016.imap',
    'observatory' : 'JWST',
    'parkey' : ('REFTYPE',),
    'sha1sum' : '0fd29b57f226f8c53bf3dfa7564eccd4e3fc56bf',
}

selector = {
    'DARK' : 'jwst_nircam_dark_0003.rmap',
    'DISTORTION' : 'N/A',
    'DRIZPARS' : 'jwst_nircam_drizpars_0000.rmap',
    'FLAT' : 'jwst_nircam_flat_0000.rmap',
    'GAIN' : 'jwst_nircam_gain_0001.rmap',
    'IPC' : 'jwst_nircam_ipc_0000.rmap',
    'LASTFRAME' : 'N/A',
    'LINEARITY' : 'jwst_nircam_linearity_0003.rmap',
    'MASK' : 'jwst_nircam_mask_0002.rmap',
    'PHOTOM' : 'jwst_nircam_photom_0002.rmap',
    'READNOISE' : 'jwst_nircam_readnoise_0001.rmap',
    'REGIONS' : 'N/A',
    'RESET' : 'N/A',
    'SATURATION' : 'jwst_nircam_saturation_0002.rmap',
    'SPECWCS' : 'N/A',
    'SUPERBIAS' : 'jwst_nircam_superbias_0003.rmap',
    'WCSREGIONS' : 'N/A',
}
\end{verbatim}
\end{mdframed}
\caption{An example of an instrument contex file (.imap)}
\label{fig:example_instrument_context}
\end{figure*}

@@

\begin{figure*}
\begin{mdframed}
\begin{verbatim}
header = {
    'derived_from' : 'jwst_0068.pmap',
    'description' : 'Hand edited to support updated .imaps for N/A',
    'mapping' : 'PIPELINE',
    'name' : 'jwst_0069.pmap',
    'observatory' : 'JWST',
    'parkey' : ('META.INSTRUMENT.NAME',),
    'sha1sum' : '691d78053ef710f31bfd8f5bea8ffdc3c1dec1aa',
}

selector = {
    'FGS' : 'jwst_fgs_0016.imap',
    'MIRI' : 'jwst_miri_0043.imap',
    'NIRCAM' : 'jwst_nircam_0016.imap',
    'NIRISS' : 'jwst_niriss_0024.imap',
    'NIRSPEC' : 'jwst_nirspec_0026.imap',
}
\end{verbatim}
\end{mdframed}
\caption{An example of a pipeline context file (.pmap)}
\label{fig:example_pipeline_context}
\end{figure*}

@@

\begin{figure*}[ht!]
\centering
\includegraphics[width=180mm]{JWST_Public_Home_Page.png}
\caption{JWST CRDS Public Home Page}
\label{fig:publichome}
\end{figure*}

@@

>>>>

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}
\usepackage{float}
\floatstyle{boxed} 
\restylefloat{figure}
%\usepackage{mdframed}
%% The amssymb package provides various useful mathematical symbols
% \usepackage{amssymb}
\usepackage{gensymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

We describe a software architecture and implementation for using rules to
determine which calibration files are appropriate for calibrating a
given observation. This new system, the Calibration Reference Data System
(CRDS), replaces what had been previously used for
the Hubble Space Telescope (HST) calibration pipelines, 
the Calibration Database System (CDBS). 
CRDS will be used for the James Webb Space
Telescope (JWST) calibration pipelines, and is currently
being used for HST calibration pipelines. CRDS can be
easily generalized for use in similar applications that need a rules-based
system for selecting the appropriate item for a given dataset; we give some
examples of such generalizations that will likely be used for JWST. The core
functionality of the Calibration Reference Data System is available under an
Open Source license. CRDS is briefly constrasted with a selected 
sampling of other similar systems used at other observatories.
\end{abstract}

@@

\item \textbf{Difficulty in undoing mistakes.}
A delivery of a file with mistaken
parameters could be difficult to undo. No reference file could be removed from
the database for various reasons.
(It is possible to address this easily by changing the schema;
nevertheless, the existing design made 
such a change difficult.)
In a number of cases, particularly if the
date that it was targeted for was incorrect, it could seriously corrupt the
selection logic. To take the most common error case: Files often had a
``useafter'' date associated with them that indicated that it should be used for
a date after the specified one \textit{if no other file had a later date that was
still before the date of the observation}. If one submitted a file intended to
be used for data after 2000-01-01 but mistakenly provided
1999-01-01 then two new file submissions are required to correct the problem.
One to resubmit the new file with the correct date; the second to resubmit the file
intended to be used after 1999-01-01 so that its useafter date is one second after
the mistaken submission.

@@

\begin{figure*}[ht!]
%%\begin{mdframed}
\begin{verbatim}
header = {
    'derived_from' : 'jwst_miri_dark_0007.rmap',
    'filekind' : 'DARK',
    'instrument' : 'MIRI',
    'mapping' : 'REFERENCE',
    'name' : 'jwst_miri_dark_0009.rmap',
    'observatory' : 'JWST',
    'parkey' : (('META.INSTRUMENT.DETECTOR','META.EXPOSURE.READPATT',
    'META.SUBARRAY.NAME'),),
    'sha1sum' : 'bd99f4960096de9ad39a9ce506d5ff985847c4a3',
}

selector = Match({
    ('MIRIFULONG', 'FAST', 'FULL') : UseAfter({
        '2015-01-27 12:05:34' 'jwst_miri_dark_0029.fits',
        '2015-03-22 00:29:01' 'jwst_miri_dark_0032.fits',
        '2015-04-07 09:14:00' 'jwst_miri_dark_0047.fits',
        ...
        }),
    ('MIRIFUSHORT', 'SLOW', 'FULL') : UseAfter({
        '2015-01-25 12:06:34' 'jwst_miri_dark_0099.fits',
        '2015-03-21 00:28:01' 'jwst_miri_dark_0101.fits',
        '2015-04-08 09:13:00' 'jwst_miri_dark_0127.fits',
        ...
        }),
        ...
})
\end{verbatim}
%%\end{mdframed}
\caption{Example showing basic parameter matching rules coupled with date dependency.}
\label{fig:example1}
\end{figure*}

@@

\begin{figure*}
%\begin{mdframed}
\begin{verbatim}
selector = SelectVersion({
   '<3.1':    'jwst_miri_flat_0065.fits',
   '<5':      'jwst_miri_flat_0073.fits',
   'default': 'jwst_miri_flat_123.fits',
})
\end{verbatim}
%\end{mdframed}
\caption{Example showing use of the software version selection mechanism.}
\label{fig:example2}
\end{figure*}

@@

\begin{figure*}
%\begin{mdframed}
\begin{verbatim}
selector = Bracket({
    1.2: "cref_flatfield_120.fits",
    1.5: "cref_flatfield_124.fits",
    5.0: "cref_flatfield_137.fits",
})
\end{verbatim}
%\end{mdframed}
\caption{Example showing bracketing selection.}
\label{fig:example3}
\end{figure*}

@@

\begin{figure*}
%\begin{mdframed}
\begin{verbatim}
selector = Match({
    ('WFC', 'N/A', 'BETWEEN 1220 1530', 'FR782N') :  "hst_wfc3_flatfile_0027.fits",
    ('WFC', 'N/A', 'BETWEEN 1530 1690', 'FR782N') :  "hst_wfc3_flatfile_0028.fits",
    ...
})
\end{verbatim}
%\end{mdframed}
\caption{An example showing matching specified ranges.}
\label{fig:example4}
\end{figure*}

@@

\begin{figure*}
%\begin{mdframed}
\begin{verbatim}
header = {
    'derived_from' : 'jwst_nircam_0015.imap',
    'instrument' : 'NIRCAM',
    'mapping' : 'INSTRUMENT',
    'name' : 'jwst_nircam_0016.imap',
    'observatory' : 'JWST',
    'parkey' : ('REFTYPE',),
    'sha1sum' : '0fd29b57f226f8c53bf3dfa7564eccd4e3fc56bf',
}

selector = {
    'DARK' : 'jwst_nircam_dark_0003.rmap',
    'DISTORTION' : 'N/A',
    'DRIZPARS' : 'jwst_nircam_drizpars_0000.rmap',
    'FLAT' : 'jwst_nircam_flat_0000.rmap',
    'GAIN' : 'jwst_nircam_gain_0001.rmap',
    'IPC' : 'jwst_nircam_ipc_0000.rmap',
    'LASTFRAME' : 'N/A',
    'LINEARITY' : 'jwst_nircam_linearity_0003.rmap',
    'MASK' : 'jwst_nircam_mask_0002.rmap',
    'PHOTOM' : 'jwst_nircam_photom_0002.rmap',
    'READNOISE' : 'jwst_nircam_readnoise_0001.rmap',
    'REGIONS' : 'N/A',
    'RESET' : 'N/A',
    'SATURATION' : 'jwst_nircam_saturation_0002.rmap',
    'SPECWCS' : 'N/A',
    'SUPERBIAS' : 'jwst_nircam_superbias_0003.rmap',
    'WCSREGIONS' : 'N/A',
}
\end{verbatim}
%\end{mdframed}
\caption{An example of an instrument contex file (.imap)}
\label{fig:example_instrument_context}
\end{figure*}

@@

\begin{figure*}
%\begin{mdframed}
\begin{verbatim}
header = {
    'derived_from' : 'jwst_0068.pmap',
    'description' : 'Hand edited to support updated .imaps for N/A',
    'mapping' : 'PIPELINE',
    'name' : 'jwst_0069.pmap',
    'observatory' : 'JWST',
    'parkey' : ('META.INSTRUMENT.NAME',),
    'sha1sum' : '691d78053ef710f31bfd8f5bea8ffdc3c1dec1aa',
}

selector = {
    'FGS' : 'jwst_fgs_0016.imap',
    'MIRI' : 'jwst_miri_0043.imap',
    'NIRCAM' : 'jwst_nircam_0016.imap',
    'NIRISS' : 'jwst_niriss_0024.imap',
    'NIRSPEC' : 'jwst_nirspec_0026.imap',
}
\end{verbatim}
%\end{mdframed}
\caption{An example of a pipeline context file (.pmap)}
\label{fig:example_pipeline_context}
\end{figure*}

@@

\floatstyle{plain}
\restylefloat{figure}
\begin{figure*}[ht!]
\centering
\includegraphics[width=180mm]{JWST_Public_Home_Page.png}
\caption{JWST CRDS Public Home Page}
\label{fig:publichome}
\end{figure*}

@@

\section{Comparison with Approaches Used at Selected Observatories}

\subsection{Chandra}

The approach Chandra uses is based on the CALDB approach developed by 
HEASARC and used by most high-engergy missions 
(\citealt{2006SPIE.6270E..1XG}). Despite
the name, no database is involved. In fact, the approach is a simpler 
version of what is described in this paper. Essentially, the rules for 
selecting files is kept in index files that list all of the available 
calibration reference files and the situations they should be used for. 
It shares some of the advantages of our approach such as being able to 
recreate past recommendations and being able to distribute the software, 
though it isn’t as simple to set up.

\subsection{Spitzer}

Spitzer uses perhaps the purest database approach we are aware of, and 
has been described in \citealt{2005ASPC..347..594L}. 
While it does permit an easy undo of updates, 
it is difficult to recreate past recommendations, and is not easy to 
distribute and run elsewhere (besides, the code and the associate 
database does not appear to be available to do so).
The rules for selection of calibration reference 
files are basically expressed as as SQL queries to the database.
Special post query processing is sometimes done, such as generating
a time-interpolated reference from two bracketing files.

\subsection{Gemini}

Gemini uses a hybrid approach, using a database to store much of the 
information, but where much of the selection of calibration reference 
files is performed by running Python code to generate the appropriate 
SQL query to apply to the database (Paul Hirst, private communication). 
The use of pure SQL queries did not capture all the needs of the 
complex selection rules required; their hybrid approach allows arbitrarily
complex rules to be constructed. It does not permit easy distribution of the 
system though it does provide a web-based service to provide calibration 
reference file selections to external users. The results rank multiple 
possible files that may be used as to their applicability. It does try 
to allow for multiple kinds of processing that may require different 
versions of calibration files depending on the kind of processing used 
(e.g., whether you have done an overscan subtraction will affect which 
bias file you are given). As such, it isn’t easy to support multiple 
versions of the rules (that requires multiple servers), nor can the 
system be easily distributed or customized without a lot of expertise 
(but it can be used remotely through the web, and there is consideration 
being given to mechanisms to allow local customization to the web results). 
Reversion is possible by using version control to reinstate the 
appropriate Python modules, but has never been used in practice.
